A decade of lost precision: Introducing a general framework for image feature matching without geometric constraints

### 2014-04-10 text/journal/paper.tex
Removed figures containing planar plots from paper

\section{Experiments and Results}
\label{S:Experiments}
%
\begin{figure}[htb]
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_A}
        \caption{Query Image}
        \label{fig:frog_a}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad a 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_B}
        \caption{Auxiliary Image}
        \label{fig:frog_b}
    \end{subfigure}%
    \\ %add desired spacing between images, e. g. ~, \quad, \qquad 
    %a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C}
        \caption{Target Image}
        \label{fig:frog_c}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C_lines}
        \caption{Target Image with epipolar constraints}
        \label{fig:frog_c_lines}
    \end{subfigure}%
    \caption{Creating epipolar constraints based on three source images.  
        In a) the \emph{query image} has been marked with the position 
        of the feature we are attempting to match. In b) the 
        \emph{auxiliary image} is shown. This image is taken with the 
        same rotation as the \emph{query image} but from a higher 
        viewpoint. The line going through the image is the epipolar line 
        of the feature point in the \emph{query image}.  The markers 
        indicate all feature points in the image found near the epipolar 
        line.  In c) we show the \emph{target image} which in this case 
        is rotated 45 degrees from the \emph{query image}.  The line 
        overlaid on the image represents the epipolar line corresponding 
        to the feature point shown in the \emph{query image}.  The 
        markers indicate all feature points in the image found near the 
    epipolar line. In d) we show the \emph{target image} again, this 
    time overlaid with the epipolar lines corresponding to all the 
    features shown in the \emph{auxiliary image}. A true correspondence 
    should be found within a small distance of one of the intersections 
    of the line in c) and the lines in d). In this particular case both 
    feature points shown in c) and d) could possible be a correct 
    correspondence.
}%
    \label{fig:frog}%
\end{figure}%
In this section we will go over the experimental results comparing 
\emph{Self-Match}, \emph{Ratio-Match}, \emph{Ratio-Match-Ext} and 
\emph{Mirror-Match}. We use two different image databases to compile 
results over different variations.  To measure results on realistic and 
variate non-planar images we test the algorithm on the 3D Objects 
database released by Moreels and Pietro \cite{moreels2007evaluation} in 
Section~\ref{S:3dobjects}. The database contains a set of 86 3D objects 
photographed from all sides at 5 degree intervals from two different 
viewpoints and under three different lighting conditions.
To measure the performance over variations such as viewpoint change, 
blur, compression noise, rotation and zoom we also use the graffiti 
image set introduced by Mikolajczyk and Schmid 
\cite{mikolajczyk2005performance} in Section~\ref{S:Planar}.  This 
dataset contains 8 scenes each with 6 images linked by homographies and 
for that reason feature only planar or almost planar scenes.

@@

\subsection{Planar Scenes}
\label{S:Planar}

\begin{figure*}[t!]
	\centering
    \includegraphics[width=2\columnwidth]{images/results_acr_all}
    \caption{Accumulated results for each scene in the image dataset by 
    Mikolajczyk and Schmid \cite{mikolajczyk2005performance}.}
    \label{fig:acr_objects}
\end{figure*}

\begin{figure}[t]
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/bark}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/bikes}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/boat}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/graf}
    \end{subfigure}%
    \vspace{1.5 mm}
    % Linebreak
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/leuven}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/trees}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/ubc}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/wall}
    \end{subfigure}%
    \caption{Scenes from the image dataset by Mikolajczyk and Schmid 
    \cite{mikolajczyk2005performance} ordered as they appear in 
Figure~\ref{fig:acr_objects}.}
    \label{fig:acr_images}
\end{figure}

To evaluate local feature descriptors, Mikolajczyk and Schmid introduced
a serie of $8$ image sets featuring different transformations such as 
compression noise, rotation and zoom, blur and viewpoint change. All 
images are tied by homographies which makes them convenient for testing 
purposes but as a consequence they all depict planar scenes which is 
restrictive in terms of image variation. Each scene contains 6 different 
images ranged in order of increasing levels of transformations. The 
transformation between image 1 and image 2 in a scene is only slight 
while image 6 for the most cases is markedly different than image 1.

For each scene we match the first image in the scene with image 3, 4, 5 
and 6 and show the accumulated result weighed by the possible number of 
true correspondences for each pair. We exclude the comparison with image 
2 because it is trivial in most cases and does not tell us much about 
the relative performance of the different methods. We find inliers by 
using the homography between image pairs as follows: Given a potential 
match between two pixels $p_1$ and $p_2$, $m = \left(p_1, p_2\right)$, 
and a homography $H$ relating the two images $I_1$ and $I_2$, we can 
calculate if $m$ is an inlier by checking if the two points satisfy the 
following criteria:
\begin{equation*}
\left\vert H p_1 - p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 \right\vert < d_{\max}
\end{equation*}
That is, the distance between $p_1$ translated to $I_2$ and $p_2$ 
\emph{plus} the distance between $p_2$ translated to $I_1$ should be 
less than a certain threshold (we use $d_{\max}=10$ pixels here).

Figure~\ref{fig:acr_objects} shows the results plotted with precision on 
the y-axis and recall along the x-axis. For scenes such as `bark, `boat' 
(zoom and rotation), `leuven' (light) and `wall' (viewpoint) all methods 
in the framework show similar results, largely because the precision for 
all methods approach 100\% for low recall which does not leave much room 
for improvement. For `bikes' and `trees' (blur) as well as `ubc'
(compression) and `graf' (viewpoint) we see \emph{Mirror-Match} clearly 
outperforming \emph{Ratio-Match} although with a smaller margin than 
with the 3D objects dataset. Because many images in the dataset are 
trivial to match and a few are very difficult we see less spread in 
between the algorithms than with the 3D objects dataset. However the 
theoretical predictions still hold true with \emph{Mirror-Match} 
performing better or equal to \emph{Ratio-Match-Ext} which in turns 
perform better or equal to \emph{Ratio-Match}. As for \emph{Self-Match} 
we are again confirmed that we can achieve performance rivaling 
\emph{Ratio-Match} using only the \emph{query image} as \emph{baseline 
set}.

@@

    \begin{subfigure}[t]{1\textwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/results_descriptors_easy}
        \caption{Keypoint / Descriptor combinations evaluated on `Bikes' 
        scene}
        \label{fig:descriptors_easy}
    \end{subfigure}%
    \caption{a) Keypoint / Descriptor combinations measured on 15 pairs 
    of photos of 3D objects taken 15 degrees apart. K = Keypoint, D = 
Descriptor\. b) Same experiment evaluated on a simple to match image 
pair; image 1 and 3 in the `Bikes' scene from Mikolaczyk and Schmid's 
review of descriptors \cite{mikolajczyk2005performance}.}
\end{figure*}

@@

To test this hypothesis we looked at the results from matching an image 
pair with very little inter image variation. We compare two image 1 and 
image 3 from the `bikes' scene in the image set released by Mikolajczyk 
and Schmid \cite{mikolajczyk2005performance}. The results are displayed 
in Figure~\ref{fig:descriptors_easy} and show a market improvement for 
\emph{Self-Match}, \emph{Mirror-Match} and \emph{Ratio-Match-Ext} over 
\emph{Ratio-Match} for the case of BRISK, BRIEF and FREAK as expected.  
For SURF however the algorithms perform similarly at low recall while 
\emph{Ratio-Match} show better performance at more lenient thresholds.

@@

\begin{table}[htb]
\caption{Complexity and running times tested on 15 image pairs on 3 
lighting conditions with average $n = 236.73$ and average $m = 247.38$ 
as tested on a Intel\textregistered\ Core\texttrademark\ i5-3550 CPU @ 
3.30~GHz with 8~GB memory.}
\label{table:running_times}
	\centering
%	\small
    \begin{tabular}{{l}{l}{l}}
    Algorithm & Complexity & Running Time\\
    \hline
    \noalign{\smallskip}
    %
    \emph{Self-Match} & $O(n\log(nm))$ & 118s  \\
    \emph{Self-Match-Ext} & $O(n\log(n(n+m)))$ & 109s\\
    \emph{Ratio-Match} & $O(n\log(m))$ & 114s\\
    \emph{Ratio-Match-Ext} & $O(n\log(n+m))$ & 112s\\
    \emph{Mirror-Match} & $O(n\log(n(n+m)))$ & 116s\\
    \emph{Mirror-Match-Ext} & $O(n\log(n+m))$ & 110s \\
    \hline
\end{tabular}
\end{table}

@@

From an evaluation using images of rotated 3D objects we have shown that
the theoretical conclusions are reflected in experimental data with 
\emph{Mirror-Match} often outperforming \emph{Ratio-Match}
significantly over 3024 image pairs. We went on to show that this 
behaviour was repeated for the classical `graffiti' imageset although 
with less variation in between the algorithms. This performance gain 
comes free in terms of both computational complexity as well as actual 
time consumption as measured over 45 image pairs. Finally evaluations 
across several different descriptors show that the gains in performance 
are tied to descriptor performance with significantly smaller gains 
occuring with descriptors that are less performant.

>>>>

\section{Experiments and Results}
\label{S:Experiments}
%
\begin{figure}[htb]
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_A}
        \caption{Query Image}
        \label{fig:frog_a}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad a 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_B}
        \caption{Auxiliary Image}
        \label{fig:frog_b}
    \end{subfigure}%
    \\ %add desired spacing between images, e. g. ~, \quad, \qquad 
    %a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C}
        \caption{Target Image}
        \label{fig:frog_c}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C_lines}
        \caption{Target Image with epipolar constraints}
        \label{fig:frog_c_lines}
    \end{subfigure}%
    \caption{Creating epipolar constraints based on three source images.  
        In a) the \emph{query image} has been marked with the position 
        of the feature we are attempting to match. In b) the 
        \emph{auxiliary image} is shown. This image is taken with the 
        same rotation as the \emph{query image} but from a higher 
        viewpoint. The line going through the image is the epipolar line 
        of the feature point in the \emph{query image}.  The markers 
        indicate all feature points in the image found near the epipolar 
        line.  In c) we show the \emph{target image} which in this case 
        is rotated 45 degrees from the \emph{query image}.  The line 
        overlaid on the image represents the epipolar line corresponding 
        to the feature point shown in the \emph{query image}.  The 
        markers indicate all feature points in the image found near the 
    epipolar line. In d) we show the \emph{target image} again, this 
    time overlaid with the epipolar lines corresponding to all the 
    features shown in the \emph{auxiliary image}. A true correspondence 
    should be found within a small distance of one of the intersections 
    of the line in c) and the lines in d). In this particular case both 
    feature points shown in c) and d) could possible be a correct 
    correspondence.
}%
    \label{fig:frog}%
\end{figure}%
In this section we will go over the experimental results comparing 
\emph{Self-Match}, \emph{Ratio-Match}, \emph{Ratio-Match-Ext} and 
\emph{Mirror-Match}. We use two different image databases to compile 
results over different variations.  To measure results on realistic and 
variate non-planar images we test the algorithm on the 3D Objects 
database released by Moreels and Pietro \cite{moreels2007evaluation} in 
Section~\ref{S:3dobjects}. The database contains a set of 86 3D objects 
photographed from all sides at 5 degree intervals from two different 
viewpoints and under three different lighting conditions.

@@

@@

    \caption{Keypoint / Descriptor combinations measured on 15 pairs of 
    photos of 3D objects taken 15 degrees apart. K = Keypoint, D = 
Descriptor\.}
\end{figure*}

@@

@@

\begin{table}[htb]
\caption{Complexity and running times tested on 15 image pairs on 3 
lighting conditions with average $n = 236.73$ and average $m = 247.38$ 
as tested on a Intel\textregistered\ Core\texttrademark\ i5-3550 CPU @ 
3.30~GHz with 8~GB memory.}
\label{table:running_times}
	\centering
%	\small
    \begin{tabular}{{l}{l}{l}}
    Algorithm & Complexity & Running Time\\
    \hline
    \noalign{\smallskip}
    %
    \emph{Self-Match} & $O(n\log(nm))$ & 118s  \\
    %\emph{Self-Match-Ext} & $O(n\log(n(n+m)))$ & 109s\\
    \emph{Ratio-Match} & $O(n\log(m))$ & 114s\\
    \emph{Ratio-Match-Ext} & $O(n\log(n+m))$ & 112s\\
    \emph{Mirror-Match} & $O(n\log(n(n+m)))$ & 116s\\
    %\emph{Mirror-Match-Ext} & $O(n\log(n+m))$ & 110s \\
    \hline
\end{tabular}
\end{table}

@@

From an evaluation using images of rotated 3D objects we have shown that
the theoretical conclusions are reflected in experimental data with 
\emph{Mirror-Match} often outperforming \emph{Ratio-Match}
significantly over 3024 image pairs. This performance gain comes free in 
terms of both computational complexity as well as actual time 
consumption as measured over 45 image pairs. Finally evaluations across 
several different descriptors show that the gains in performance are 
tied to descriptor performance with significantly smaller gains occuring 
with descriptors that are less performant.
